---
title: "homework 07"
author: "Yifan Zhang"
date: "October 26, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth",dev="png",fig.align  = 'center')
pacman::p_load(
ggplot2,
knitr,
arm,
data.table,
foreign,
gridExtra,
car,
stringr,
rstan,
rstanarm,
zoo, 
dplyr, 
tidyverse, 
magrittr, 
reshape2,
stringr, 
merTools
)
```



# Data analysis 


## CD4 percentages for HIV infected kids

The folder `cd4` has CD4 percentages for a set of young children with HIV who were measured several times over a period of two years. The dataset also includes the ages of the children at each measurement.

```{r,echo=FALSE}
# Read in the data from an excel-format ".csv" file
hiv.data.raw <- fread ("http://www.stat.columbia.edu/~gelman/arm/examples/cd4/allvar.csv")

invisible(hiv.data.raw[,ok := !is.na(CD4PCT) ])

hiv.data<-hiv.data.raw[ok==TRUE]
invisible(hiv.data[,y :=sqrt (CD4PCT)])
 # kid's age (yrs) at the beginning of the study
invisible(hiv.data[,age.baseline := baseage ]  ) 
# kids age (yrs) at the time of measurement
invisible(hiv.data[,age.measurement := visage ] ) 
invisible(hiv.data[,time := visage - baseage ] )
setnames(hiv.data,"treatmnt","treatment") 
hiv.data<-hiv.data[complete.cases(hiv.data[,list(y,time,age.baseline,treatment)])]
```

1. Graph the outcome (the CD4 percentage, on the square root scale) for each child as a function of time.
```{r}
data.cd4 <- hiv.data %>% dplyr::select(newpid, treatment, CD4PCT, y: time)
length(unique(hiv.data$newpid))
cd4_plot <- ggplot(hiv.data %>% dplyr::select(newpid,y: time), aes(x = time, y = y)) +
  labs(y = "square root of CD4 percentage", x = "Time(visit age - baseline age)") + 
  theme(legend.position = "none")
cd4_plot + geom_line() + geom_point() + aes(color = factor(newpid), group = factor(newpid))

# down to show 5% of children colored and 95% of children greyed
sample <- sample(unique(hiv.data$newpid), round(length(unique(hiv.data$newpid))*.05), replace = FALSE)
cd4_plot.1 <-  ggplot(data = hiv.data, aes(x = time, y = y, group = factor(newpid)),) + 
  geom_line(color = "grey", lwd = .5) + geom_point(color = "grey", size = .5) + theme_bw() + 
  labs(y = "square root of CD4 percentage", x = "Time(visit age - baseline age)")

cd4_plot.1 + geom_line(data = hiv.data[hiv.data$newpid %in% sample, ], mapping = aes(x = time, y = y, group = factor(newpid), color = factor(newpid)), lwd = 1) +
  geom_point(data = hiv.data[hiv.data$newpid %in% sample, ], mapping = aes(x = time, y = y, group = factor(newpid), color = factor(newpid)), size = .8)

```

2. Each child's data has a time course that can be summarized by a linear fit. Estimate these lines and plot them for all the children.

```{r}
ggplot(data.cd4, aes(x = time, y = y, group = factor(newpid), color = factor(newpid))) + 
  geom_smooth(se = FALSE, method = "lm", lwd = .3, formula = 'y ~ x') + theme(legend.position = "none") +
  labs(x = "Time(visit age - baseline age)", y = "square root of CD4 percentage")
```


3. Set up a model for the children's slopes and intercepts as a function of the treatment and age at baseline. Estimate this model using the two-step procedure–first estimate the intercept and slope separately for each child, then fit the between-child models using the point estimates from the first step.
```{r}
# first estimate the intercept and slope separately for each child
uid <- unique(data.cd4$newpid)
regression.matrix <- data.frame(intercept = NA, slope = NA)
for(i in 1: length(uid)){
  regression.matrix[i, ] <- coef(lm(y ~ time, data = subset(data.cd4, newpid == uid[i])))
}
regression.matrix[is.na(regression.matrix)] <- 0
regression.matrix <- data.frame(newpid = uid, regression.matrix); rownames(regression.matrix) <- c()
# fit the between-child models
u.data.cd4 <- unique(data.cd4 %>% dplyr::select(newpid, treatment, age.baseline))
regression.matrix <- left_join(regression.matrix, u.data.cd4, by = "newpid")
fit_intercept <- lm(intercept ~ factor(treatment) + age.baseline, data = regression.matrix)
fit.slope     <- lm(slope     ~ factor(treatment) + age.baseline, data = regression.matrix)
summary(fit_intercept)
summary(fit.slope)
```


4. Write a model predicting CD4 percentage as a function of time with varying intercepts across children. Fit using `lmer()` and interpret the coefficient for time.
```{r}
lmer.fit.1 <- lmer(CD4PCT ~ time + (1|newpid), data = data.cd4)
stan_lmer.fit.1 <- stan_lmer(CD4PCT ~ time + (1|newpid), data = data.cd4)
summary(lmer.fit.1)
# ranef(lmer.fit.1)
```


5. Extend the model in (4) to include child-level predictors (that is, group-level predictors) for treatment and age at baseline. Fit using `lmer()` and interpret the coefficients on time, treatment, and age at baseline.

```{r}
lmer.fit.2 <- lmer(CD4PCT ~ time + treatment + age.baseline + (1|newpid), data = data.cd4)
summary(lmer.fit.2)
```

6. Investigate the change in partial pooling from (4) to (5) both graphically and numerically.

```{r}
# numerically comparsion
coef.lmer.fit.1 <- coef(lmer.fit.1)[["newpid"]] %>% mutate(time_slope = time)
coef.lmer.fit.1$intercept <- coef.lmer.fit.1$`(Intercept)`
coef.lmer.fit.1 %<>% dplyr::select(intercept, time_slope) 
coef.lmer.fit.2 <- coef(lmer.fit.2)[["newpid"]] %>% mutate(time_slope = time, 
                                                           treat_slope = treatment, 
                                                           baseline_slope = age.baseline)
coef.lmer.fit.2$intercept <- coef.lmer.fit.2$`(Intercept)`
coef.lmer.fit.2 %<>% dplyr::select(intercept, time_slope, treat_slope, baseline_slope) 
library(merTools)
# compare the prediction difference
pre.sample <- sample(length(data.cd4$newpid), round(.05*length(data.cd4$newpid)), replace = FALSE)
pred.lmer.fit.1 <- predictInterval(lmer.fit.1) %>% 
  add_column(model = "lmer.fit.1", number = 1:length(data.cd4$newpid))
pred.lmer.fit.2 <- predictInterval(lmer.fit.2) %>% 
  add_column(model = "lmer.fit.2", number = 1:length(data.cd4$newpid))
pred.lmer.fit   <- rbind(pred.lmer.fit.1[pre.sample, ], pred.lmer.fit.2[pre.sample, ])
ggplot(pred.lmer.fit) + 
  geom_pointrange(aes(x = number, y = fit, ymin = lwr, ymax = upr, 
                      group = factor(model), color = factor(model)))
# compare differences of random effects on intercept
plotREsim(REsim(lmer.fit.1))
plotREsim(REsim(lmer.fit.2))
```


7. Use the model fit from (5) to generate simulation of predicted CD4 percentages for each child in the dataset at a hypothetical next time point.

```{r}
stan_lmer.fit.2 <- stan_lmer(CD4PCT ~ time + factor(treatment) + age.baseline + (1|newpid), data = data.cd4)
# select hypothetical equal to 1
data.cd4.hypo <- data.cd4 %>% mutate(hypo.time = 1) %>% dplyr::select(newpid, CD4PCT, hypo.time, treatment, age.baseline)
colnames(data.cd4.hypo)[3] <- c("time")
sim.data.cd4.hypo <- posterior_predict(stan_lmer.fit.2, newdata = data.cd4.hypo, draws = 1000)
pp_check(stan_lmer.fit.2)
```

8. Use the same model fit to generate simulations of CD4 percentages at each of the time periods for a new child who was 4 years old at baseline.

```{r}
sim.data.4 <- data.cd4 %>% 
  mutate(agebase4 = 4, agemeasure = 4 + time) %>% #baseline age = 4 years old
  dplyr::select(newpid, CD4PCT, time, treatment, agebase4)
colnames(sim.data.4)[5] <- c("age.baseline")
sim.data.4 <- posterior_predict(stan_lmer.fit.2, newdata = sim.data.4, draws = 1000)
```


9. Posterior predictive checking: continuing the previous exercise, use the fitted model from (5) to simulate a new dataset of CD4 percentages (with the same sample size and ages of the original dataset) for the final time point of the study, and record the average CD4 percentage in this sample. Repeat this process 1000 times and compare the simulated distribution to the observed CD4 percentage at the final time point for the actual data.
```{r}
# create the simulate data(for the final time point of the study)
sim.number <- 100
times <- 1000
final.point.dect <- data.cd4 %>% group_by(newpid) %>% summarise(time = max(time))
data.final.point <- inner_join(data.cd4, final.point.dect, by = c("newpid", "time")) %>%
  dplyr::select(newpid, treatment, time, age.baseline, CD4PCT)
sim.matrix <- data.frame(newpid = unique(data.cd4$newpid))
for(i in 1:times){
  sim.final.point <- posterior_predict(stan_lmer.fit.2, newdata = data.final.point, 
                                       draws = sim.number) %>% apply(2, mean)
  sim.matrix[, (i+1)] <- sim.final.point
  colnames(sim.matrix)[i+1] <- paste0("iter_", i)
}
sim.matrix <- sim.matrix %>% melt(id.vars = "newpid", variable.name = "iteration", value.name = "CD4PCT")

# create pp_check
pp_check.final.point <- ggplot() + 
  geom_density(data = sim.matrix, aes(x = CD4PCT, group = iteration), color = "azure3", lwd = .3) +
  geom_density(data = data.final.point, mapping = aes(CD4PCT), color = "red", lwd = 1) + theme_bw()
pp_check.final.point
# not really good fitness for final points! 
```

10. Extend the model to allow for varying slopes for the time predictor.
```{r}
lmer.fit.3 <- lmer(CD4PCT ~ time + treatment + age.baseline + (1 + time|newpid), data = data.cd4)
stan_lmer.fit.3 <- stan_lmer(CD4PCT ~ time + treatment + age.baseline + (1 + time|newpid), data = data.cd4)
summary(lmer.fit.3)
```


11. Next fit a model that does not allow for varying slopes but does allow for different coefficients for each time point (rather than fitting the linear trend).
```{r}
lmer.fit.4 <- lmer(CD4PCT ~ time + treatment + age.baseline + (1|time), data = data.cd4)
stan_lmer.fit.4 <- stan_lmer(CD4PCT ~ time + treatment + age.baseline + (1|time), data = data.cd4)
summary(lmer.fit.4)
```


12. Compare the results of these models both numerically and graphically.
```{r}
# up till now, we get four linear mixed effect models
lmer.fit.1
lmer.fit.2
lmer.fit.3
lmer.fit.4
# numerically 
anova(lmer.fit.1, lmer.fit.2, lmer.fit.3, lmer.fit.4)
AIC(lmer.fit.1, lmer.fit.2, lmer.fit.3, lmer.fit.4)
# from the point of view of anova and AIC test, lmer.fit.3 looks better

# graphically 
pp_check(stan_lmer.fit.1)
pp_check(stan_lmer.fit.2)
pp_check(stan_lmer.fit.3)
pp_check(stan_lmer.fit.4)
```


## Figure skate in the 1932 Winter Olympics

The folder olympics has seven judges' ratings of seven figure skaters (on two criteria: "technical merit" and "artistic impression") from the 1932 Winter Olympics. Take a look at 
http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt
```{r,echo=FALSE}
filename<- "http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt"
olympics1932_na<-read.fwf(filename,widths=c(2,14,9,9,9,9,9,9,9),skip=21,header = FALSE)
colnames(olympics1932_na)<- c("pair", "criterion", "judge_1",  "judge_2",  "judge_3",
                              "judge_4",  "judge_5" , "judge_6",  "judge_7")

olympics1932<-na.locf(olympics1932_na)
olympics1932$criterion<-str_trim(olympics1932_na$criterion)
```

1. Construct a $7\times 7 \times 2$ array of the data (ordered by skater, judge, and judging criterion).

```{r}
olympics1932_program     <- olympics1932 %>% subset(criterion == "Program")
olympics1932_performance <- olympics1932 %>% subset(criterion == "Performance")
olympics <- list(olympics1932_program, olympics1932_performance)
olympics[[2]]
```

2. Reformulate the data as a $98\times 4$ array (similar to the top table in Figure 11.7), where the first two columns are the technical merit and artistic impression scores, the third column is a skater ID, and the fourth column is a judge ID.

```{r}
olympics1932.data <- olympics1932 %>% 
  melt(id = c("pair", "criterion"), variable.name = "judge_ID", value.name = "scores") %>%
  dplyr::select(c("criterion", "scores", "pair", "judge_ID"))
```

3. Add another column to this matrix representing an indicator variable that equals 1 if the skater and judge are from the same country, or 0 otherwise.

```{r}
judge <- data.frame(judge_ID = unique(olympics1932.data$judge_ID), judge = c("Jeno Minich, Hungary", "Yngvar Bryn, Norway", "Hans Grunauer, Austria", "Walter Jakobsson, Finland", "George Torchon, France", "Herbert J. Clarke, Great Britain", "Charles M. Rotch, United States"))
skater <- data.frame(pair = unique(olympics1932.data$pair), skater = c("Andree Brunet, Pierre Brunet, France", "Beatrix Loughran, Sherwin Badger, United States", "Emilia Rotter, Laszlo Szollas, Hungary", "Olva Oronista, Sandor Szalay, Hungary","Constance Wilson-Samuel, Montgomery Wilson, Canada", "Frances Claudet, Chauncey Bangs, Canada", "Gertrude Meredith, Joseph K. Savage, United States"))
olympics1932.data <- inner_join(olympics1932.data, judge, by = "judge_ID")
olympics1932.data <- inner_join(olympics1932.data, skater, by = "pair")
olympics1932.data <- olympics1932.data %>% 
  separate(judge, c("judge_name", "judge_country"), sep = ", ", fill = "right")
skater_info <- str_split_fixed(string = olympics1932.data$skater, pattern = ", ", n = 3)
olympics1932.data <- olympics1932.data %>% mutate(skater_1 = skater_info[, 1], 
                                                  skater2 = skater_info[, 2], 
                                                  skater_country = skater_info[, 3])
olympics1932.data <- olympics1932.data %>% 
  mutate(indicator = ifelse(olympics1932.data$judge_country == olympics1932.data$skater_country, 1, 0)) %>% dplyr::select(criterion, scores, pair, judge_ID, judge_country, skater_country, indicator)
```

4. Write the notation for a non-nested multilevel model (varying across skaters and judges) for the technical merit ratings and fit using lmer().
```{r}
data_4 <- olympics1932.data %>% subset(criterion == "Performance")
fit_program <- lmer(scores ~ 1 + (1|pair) + (1|judge_ID), data = data_4)
fixef(fit_program)
ranef(fit_program)$pair
ranef(fit_program)$judge_ID
```

5. Fit the model in (4) using the artistic impression ratings.
```{r}
data_5 <- olympics1932.data %>% subset(criterion == "Program")
fit_performance<- lmer(scores ~ 1 + (1|pair) + (1|judge_ID), data = data_5)
fixef(fit_performance)
ranef(fit_performance)$pair
ranef(fit_performance)$judge_ID
```

6. Display your results for both outcomes graphically.

```{r}
data_4 <- data_4 %>% mutate(prediction = predict(fit_program, newdata = data_4))
data_5 <- data_5 %>% mutate(prediction = predict(fit_performance, newdata = data_5))
plot_data.6 <- rbind(data_4, data_5) %>% mutate(judge = as.numeric(plot_data.6$judge_ID))
ggplot(plot_data.6, aes(x = judge, y = prediction, group = interaction(pair, criterion), color = factor(pair))) + geom_smooth(aes(linetype = criterion), se = FALSE, formula = 'y ~ x', method = 'loess', lwd = .5)
```

7. (optional) Use posterior predictive checks to investigate model fit in (4) and (5).

```{r}
stan_program     <- stan_lmer(scores ~ 1 + (1|pair) + (1|judge_ID), data = data_4)
stan_performance <- stan_lmer(scores ~ 1 + (1|pair) + (1|judge_ID), data = data_5)
sim_program     <- posterior_predict(stan_program, newdata = data_4, draws = 1000)
sim_performance <- posterior_predict(stan_performance, newdata = data_5, draws = 1000)
pp_check(stan_program)
pp_check(stan_performance)
```

## Different ways to write the model:

Using any data that are appropriate for a multilevel model, write the model in the five ways discussed in Section 12.5 of Gelman and Hill.

model: lmer.fit.2 <- lmer(CD4PCT ~ time + treatment + age.baseline + (1|newpid), data = data.cd4)
fixed effect: $$CD4 percentages = \beta_0 + \beta_1 \times time + \beta_2 times treatment + \beta_3 \times age.baseline$$
1. $$ $$
2. 
3. 
4. 
5. 

## Models for adjusting individual ratings: 

A committee of 10 persons is evaluating 100 job applications. Each person on the committee reads 30 applications (structured so that each application is read by three people) and gives each a numerical rating between 1 and 10.


1. It would be natural to rate the applications based on their combined scores; however, there is a worry that different raters use different standards, and we would like to correct for this. Set up a model for the ratings (with parameters for the applicants and the raters).

$S_{i1[j]}$: first score of application i, which is graded by rater j; similarly, for this application, we have $S_{i2[j]}$ and              $S_{i3[j]}$
$C_i$: combined score for application i; 

the model I defined is: $$ C_i = S_{i1[j]} + S_{i2[j]} + S_{i3[j]} + (1|G_1) + (1|G_2) + (1|G_3)$$ and $G_1$, $G_2$ and $G_3$ are the group of application i 's first, second and third rate. then, the next step is to estimated random effects of different rater. 
for rater j, he/she rates $S_{k_1[j]}, \dots, S_{k_{30}[j]}$; for each rater, the applications he/she rate randomly come from 100 applications and should follow same distribution. Thus, it sounds reasonable to define random effects of rater by standardized the distribution of applications rated by this rater. let $mean_j$ and $sd_j$ be the mean and standard deviation of $S_{k_1[j]}, \dots, S_{k_{30}[j]}$. set 5 as a standard center, random effect of $G_{1[j]} = G_{2[j]} = G_{3[j]} = 5 - mean_j$


2. It is possible that some persons on the committee show more variation than others in their ratings. Expand your model to allow for this.

make an adjustment to random effect and take variance into consideration. $$ C_i = S_{i1[j]} + S_{i2[j]} + S_{i3[j]} + (1 +S_{i1[j]} |G_1) + (1 + S_{i2[j]}|G_2) + (1 + S_{i3[j]}|G_3)$$. set 5 as a standard center and 1 as the standardized sd, random effect of $G_{1[j]} = G_{2[j]} = G_{3[j]}$ on intercept is $(5 - mean_j)/sd_j$ and random effect of $G_{1[j]} = G_{2[j]} = G_{3[j]}$ on slope is $ \frac{1}{sd_j}$.



##  Multilevel logistic regression 

The folder `speed.dating` contains data from an experiment on a few hundred students that randomly assigned each participant to 10 short dates with participants of the opposite sex (Fisman et al., 2006). For each date, each person recorded several subjective numerical ratings of the other person (attractiveness, compatibility, and some other characteristics) and also wrote down whether he or she would like to meet the other person again. Label $y_{ij} = 1$ if person $i$ is interested in seeing person $j$ again $0$ otherwise.
And $r_{ij1},\dots, r_{ij6}$ as person $i$'s numerical ratings of person $j$ on the dimensions of attractiveness, compatibility, and so forth.
Please look at 
http://www.stat.columbia.edu/~gelman/arm/examples/speed.dating/Speed%20Dating%20Data%20Key.doc
for details.

```{r}
# dating<-fread("http://www.stat.columbia.edu/~gelman/arm/examples/speed.dating/Speed%20Dating%20Data.csv")
# dating<-fread("Speed Dating Data.csv")
dating <- read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/speed.dating/Speed%20Dating%20Data.csv", header = TRUE)
```

1. Fit a classical logistic regression predicting $Pr(y_{ij} = 1)$ given person $i$'s 6 ratings of person $j$. Discuss the importance of attractiveness, compatibility, and so forth in this predictive model.

```{r}
dating <- dating %>% dplyr::select(c(iid, pid, match, attr_o, sinc_o, intel_o, fun_o, amb_o, shar_o))
dating %>% count(iid)
dating %>% count(pid)
# match		1=yes, 0=no
# attr_o: 		rating by partner the night of the event, for all 6 attributes
dating_complete_pool <- glm(match ~ attr_o + sinc_o + intel_o + fun_o + amb_o + shar_o, data=dating,                                 family = binomial)
summary(dating_complete_pool)
```

2. Expand this model to allow varying intercepts for the persons making the evaluation; that is, some people are more likely than others to want to meet someone again. Discuss the fitted model.

```{r}
dating_pooled_1 <- glmer(match~gender + attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o+(1|iid),                                    data=dating, family=binomial)
# match		1=yes, 0=no
# gender: female = 0, male = 1

# # fixed effect and random effect
# fixef(dating_pooled_1)
# ranef(dating_pooled_1)
plotREsim(REsim(dating_pooled_1))
```
discussion: factor gender(male), 

3. Expand further to allow varying intercepts for the persons being rated. Discuss the fitted model.

```{r}
dating_pooled_2 <- glmer(match ~ gender + attr_o + sinc_o + intel_o + fun_o + amb_o + shar_o + (1|iid) +                          (1|pid), data=dating,family=binomial)
fixef(dating_pooled_2)
ranef(dating_pooled_2)$iid
ranef(dating_pooled_2)$pid
plotREsim(REsim(dating_pooled_2))
```

4. You will now fit some models that allow the coefficients for attractiveness, compatibility, and the other attributes to vary by person. Fit a no-pooling model: for each person i, fit a logistic regression to the data $y_{ij}$ for the 10 persons j whom he or she rated, using as predictors the 6 ratings $r_{ij1},\dots,r_{ij6}$ . (Hint: with 10 data points and 6 predictors, this model is difficult to fit. You will need to simplify it in some way to get reasonable fits.)
```{r}
uiid <- unique(dating$iid)
dating_no_pool_list <- vector("list",length(uiid))
# for(i in 1:length(uiid)){
# #  attr_o +sinc_o +intel_o +fun_o +amb_o+shar_o,
# dating_no_pool_list[[i]] <- summary(glm(match~attr_o+shar_o,
#                        data=dating,
#                        subset = subset(dating, iid==uiid[i]),
#                        family=binomial))$coefficients
# }
# turn to use multilevel model! 
dating_partial_pooled_1 <- glmer(match ~ attr_o  + shar_o + (1 + attr_o  + shar_o|iid), data = dating, family = binomial)
fixef(dating_partial_pooled_1)
ranef(dating_partial_pooled_1)$iid
```

5. Fit a multilevel model, allowing the intercept and the coefficients for the 6 ratings to vary by the rater i.

```{r}
dating_pooled_3 <- stan_glmer(match~gender + attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o+(1+attr_o +sinc_o +intel_o +fun_o +amb_o +shar_o|iid)+(1|pid),data=as.data.frame(dating),family=binomial)
```

6. Compare the inferences from the multilevel model in (5) to the no-pooling model in (4) and the complete-pooling model from part (1) of the previous exercise.
```{r}

```

